### 2021-07-13 ch/1-introduction.tex
Add some pictures

>>>>

\Figref{fig:dataset-bias-overview} shows a visual representation of the two kinds of dataset bias discussed here.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figures/dataset_bias_overview_vertical.pdf}
  \caption{%
    Schematic representation of label bias and sampling bias.
    \textsc{Top}: the label bias changes positive labels to negative labels for \(s=0\),
    and in the opposite direction for \(s=1\) (though this not always the case).
    \textsc{Bottom}: the sampling bias \emph{intercepts} samples with a positive label from \(s=0\).
  }%
  \label{fig:dataset-bias-overview}
\end{figure}

### 2021-07-13 ch/2-related-work.tex
Add some pictures

\citet{kim2019learning} also consider the problem of learning from biased data (and evaluating on \emph{un}biased data).
While they do not formulate the problem as a fairness problem,
it is equivalent to enforcing fairness in the presence of severe sampling bias.
Their main motivating example is the Coloured MNIST dataset;
a dataset derived from the MNIST dataset~\citep{lecun1994mnist}.
In this dataset, digits are randomly coloured in the test set,
but in the training set there is a one-to-one correspondence between digit class and colour.
A na\"ive classifier will learn to predict colour instead of digit class.
The proposed method is very similar to \citet{ganin2016domain} and \citet{edwards2016censoring},
if we think of the colours as forming different domains
(\ie, the \emph{red} domain contains all red digits, etc.).
The goal is then to learn a domain-independent representation.
The main difference to previous work is an additional entropy loss term in the adversarial loss.

@@

As the authors point out,
the goal of machine learning should be to identify natural, deep, robust structures in reality,
instead of relying on superficial correlations.
(This dichotomy is sometimes framed as \emph{correlation} vs \emph{causality},
where ``causality'' takes on a very broad meaning that encompasses any kind of fundamental structure in reality.)
Under this view, methods for invariance learning can be framed as trying to learn a more fundamental structure
than the dataset at first seems to show.
By aiming to be invariant to specific environments/subgroups/domains (denoted by the variable \(s\)),
they can be seen as making the claim that \(s\) does not represent a natural structure,
but is merely an artefact of the data generation process.
Furthermore, the ground-truth-centric view of dataset bias can be understood as
aiming to identify the fundamental structures and aiming to be invariant to everything else.
This also fits with \citet{kim2019learning}'s work on Coloured MNIST:
colour is not regarded as the fundamental structure in the data.

>>>>

\begin{figure}[tp]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/cmnist-train.png}
    \caption{Training set: one-to-one mapping between colour and digit class.}%
    \label{fig:literature-cmnist-train}
  \end{subfigure}
  \quad\quad
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figures/cmnist-test.png}
    \caption{Test set: random assignment of colours.}%
    \label{fig:literature-cmnist-test}
  \end{subfigure}
  \caption{%
    A typical example of the Coloured MNIST dataset.
  }%
  \label{fig:literature-cmnist}
\end{figure}

\citet{kim2019learning} also consider the problem of learning from biased data (and evaluating on \emph{un}biased data).
While they do not formulate the problem as a fairness problem,
it is equivalent to enforcing fairness in the presence of severe sampling bias.
Their main motivating example is the Coloured MNIST dataset;
a dataset derived from the MNIST dataset~\citep{lecun1994mnist}.
In this dataset, digits are randomly coloured in the test set,
but in the training set there is a one-to-one correspondence between digit class and colour.
(See \figref{fig:literature-cmnist} for an example of this dataset.)
A na\"ive classifier will learn to predict colour instead of digit class.
The proposed method is very similar to \citet{ganin2016domain} and \citet{edwards2016censoring},
if we think of the colours as forming different domains
(\ie, the \emph{red} domain contains all red digits, etc.).
The goal is then to learn a domain-independent representation.
The main difference to previous work is an additional entropy loss term in the adversarial loss.

@@

As the authors point out,
the goal of machine learning should be to identify natural, deep, robust structures in reality,
instead of relying on superficial correlations.
(This dichotomy is sometimes framed as \emph{correlation} vs \emph{causality},
where ``causality'' takes on a very broad meaning that encompasses any kind of fundamental structure in reality; see also the discussion of \emph{causality} in \chapref{ch:introduction}.)
Under this view, methods for invariance learning can be framed as trying to learn a more fundamental structure
than the dataset at first seems to show.
By aiming to be invariant to specific environments/subgroups/domains (denoted by the variable \(s\)),
they can be seen as making the claim that \(s\) does not represent a natural structure,
but is merely an artefact of the data generation process.
Furthermore, the ground-truth-centric view of dataset bias can be understood as
aiming to identify the fundamental structures and aiming to be invariant to everything else.
This also fits with \citet{kim2019learning}'s work on Coloured MNIST:
colour is not regarded as the fundamental structure in the data.

### 2021-07-13 ch/3-structure.tex
Add some pictures

@@

In order to construct the target labels,
we use side information about summary statistics for a balanced training set.
This allows us to target a specific balanced set, instead of just any balanced set.
In other words, rather than just enforcing \ac{DP}, the method gives control over the target rates \(P(\hat{y}=1|s)\),
which are the only fairness-hyperparameters of the model.
The target labels \(\bar{y}\) represent an uncertain estimate of labels corresponding to a balanced dataset
(one where \(\bar{y} \perp s\)).
The idea is then the learn a predictor for these target labels instead of the given (biased) training labels \(y\).
Via the sum rule of probabilities, it is possible to express the model likelihood in terms of the target labels,
such that maximising the likelihood corresponds to improving the prediction of the target labels.

@@

The described method works particularly well with \acp{INN},
as they ensure that no information is lost that is unrelated to \(s\).
However, the price to pay for using \acp{INN} is higher memory requirement and slower training.
Thus, we also present a variant of the method using \iac{VAE},
which does not have the guarantee about preserving information,
but also does not suffer from the increased training cost as much.
\acp{VAE} are similar to \acp{INN}
in that their encoding conforms to a specific probability distribution,
from which we can sample our null-samples.
The choice between the two presented variants comes down
to whether the user is willing to trade higher training costs for higher-quality results.
However, the key element is any kind encoder -- producing a split-encoding --
whose output can be subjected to adversarial training,
and other kinds of encoders will potentially work as well.

>>>>

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.9\textwidth]{figures/label_bias.pdf}
  \caption{%
    A diagram of a simple case of label bias, where both the class label \(y\) and the sensitive attribute \(s\) are binary.
    On the left, we have the ideal (possibly fictional) dataset with the target labels \(\bar{y}\),
    where the proportion of positive outcomes is the same for both demographic groups;
    and on the right, the training labels, for which the proportions are \emph{not} the same.
  }%
  \label{fig:label-bias-overview}
\end{figure}

@@

In order to construct the target labels,
we use side information about summary statistics for a balanced training set.
This allows us to target a specific balanced set, instead of just any balanced set.
In other words, rather than just enforcing \ac{DP}, the method gives control over the target rates \(P(\hat{y}=1|s)\),
which are the only fairness-hyperparameters of the model.
The target labels \(\bar{y}\) represent an uncertain estimate of labels corresponding to a balanced dataset
(one where \(\bar{y} \perp s\); see also \figref{fig:label-bias-overview}).
The idea is then the learn a predictor for these target labels instead of the given (biased) training labels \(y\).
Via the sum rule of probabilities, it is possible to express the model likelihood in terms of the target labels,
such that maximising the likelihood corresponds to improving the prediction of the target labels.

@@

The described method works particularly well with \acp{INN},
as they ensure that no information is lost that is unrelated to \(s\).
However, the price to pay for using \acp{INN} is higher memory requirement and slower training.
Thus, we also present a variant of the method using \iac{VAE},
which does not have the guarantee about preserving information,
but also does not suffer from the increased training cost as much.
\acp{VAE} are similar to \acp{INN}
in that their encoding conforms to a specific probability distribution,
from which we can sample our null-samples.
The choice between the two presented variants is determined
by whether the user is willing to accept higher training costs
for a lower probability of losing information needed for any prediction tasks.
However, the key element is simply any kind encoder -- producing a split-encoding --
whose output can be subjected to adversarial training,
and other kinds of encoders will potentially work as well.
